---
title: "Robust covariance matrix under cellwise contamination"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  message = F, 
  warning = F , 
  comment = "#>"
)
```

```{r echo=FALSE}
library(cellWise)
library(snipEM)
#remotes::install_github('cran/snipEM')
#set.seed(2)
dat = robcovcell::genevar(e = 0.1)
x = dat$x
bi = dat$bi
group = dat$group
sigma = dat$sigma
```


<<<<<<< HEAD
```{r echo=FALSE, eval=FALSE}
V = matrix(1, nrow = 100, ncol = 15)
=======
```{r echo=FALSE, eval=F}
V = matrix(1, nrow = 100, ncol = 20)
>>>>>>> 4d8e60bf7b79e477e483fdd2121c73dd999322e9
V[!is.na(match(as.vector(x),boxplot(as.vector(x),plot=FALSE)$out))] = 0

fit = snipEM::snipEM(X = x, V = V)
diag(fit$S)

fit = covBC(x, group = c(rep(1,5), rep(2,5), rep(3,5)))
diag(fit$covmatrix)

```

# Methodology

Given a bunch of variables, sometimes they have different outlying types. As shown in Figure below, we'd like to divide every five columns into one group. A few cells are detected as outliers in the first group because they are clean variables. In the second group, some rows are detected as oultiers entirely. In the thrid group, some rows are outlying as well but not same as the second group. Many cellwise outliers are detected in the fourth group.

A classic technique to estimate covariance matrix (under cellwise contamination) is Orthogonalized Gnanadesikan-Kettenring (OGK) Covariance Matrix estimator (Maronna and Zamar, 2002), the main idea of which is pairwise estimation. However, under knowing group labels, we could just estimate covariance matrix by grouped cellwise weights. 

A second approach is the snipEM procedure proposed by Farcomeni (2014). Its first step flags cellwise outliers in each variable separately regard them as snipped cells (missing values). The second step tries many interchanges that unsnip a randomly chosen snipped cell and at the same time snip a randomly chosen unsnipped cell, and only keeps an interchange when it increases the partial Gaussian likelihood.

We'd like to explore two new approaches to obtain robust estimation of the covariance matrix.

The first approach is based on outlier detection. We'd like to detect outliers in each variables seperately as the first step. For simplicity here we use a boxplot rule directly. As an alternative, we could also use Minimum Covariance Determinant (MCD) estimator (Rousseeuw and van Driessen, 1999) to get the scatter and location first and then compute weights of the observations using the final estimate of the location and scatter. 

After obtaining a weight matrix $\mathbf w$, we could compute a centralized design matrix
$$
{\mathbf x_c}  = {\mathbf x} - {\mathbf 1_n}{ \hat{\mathbf\mu}_p}^\top,
$$
where ${\hat{\mathbf\mu}_p} = {\mathbf 1}_n^\top (\mathbf w \otimes\mathbf x)/{\mathbf 1}_n^\top\mathbf w$. Then, the covariance matrix could be estimated by $\mathrm{Cov} = (\mathbf w \otimes\mathbf x_c)^\top(\mathbf w \otimes\mathbf x_c)/n$.


The second approach is based on blocking and conquering methods. First, we use cellwise-robust estimators (pairwise estimator) to obtain an initial estimation of the covariance matrix. Then, for variables detected as clean columns or having rowwise outliers, we use rowwise-robust estimators (MCD estimator) to obtain a robust estimation of the covariance matrix of those variables and replace it with the corresponding block in the initial estimation. Finally, for variables detected as clean columns, we use sample covariance matrix as an estimation. 

For simplicity, here we assume that we know outlying types of all variables.


```{r echo=FALSE}

fit1 <- DDC(x[1:20,])
cellMap(fit1$remX, fit1$stdResid, columnlabels = c(1,rep(" ",13),15), rowlabels = c(1,rep(" ",18),20), columnangle = 0,
        rowtitle = "Observations", columntitle = "Variables", sizetitles = 2,adjustrowlabels = 0.5, adjustcolumnlabels = 0.5)
```

# Simulation
In our simulations, we set $n = 100$, $p = 15$, $\mathbf{x}_i$ is sampled from $N(\pmb 10,\pmb\Sigma_{xx})$. The correlation structure among variables is given by $\Sigma _{ij} = \rho^{|i-j|}$ and we set $\rho = 0.5$. 

Contamination proportions are set as $2\%$, $5\%$ and $10\%$ for all variables separately. Outlying cells of all variables are randomly generated from $0.5 N(\gamma, 1)+ 0.5N(-\gamma, 1)$ and we set 4 different kinds of outlying types. We vary $\gamma$ over the set $\{0,2,4,6,8,10\}$ to simulate outliers with different magnitudes. When we apply appropriate $\gamma$'s, those cells will be outlying cells but not not necessarily marginal outliers.

We compared 5 different methods: Minimum Covariance Determinant (MCD) estimator (Rousseeuw and van Driessen, 1999), Orthogonalized Gnanadesikan-Kettenring (OGK) Covariance Matrix estimator (Maronna and Zamar, 2002), GK pairwise estimator (Gnanadesikan and Kettenring, 1972), cellwise weighted estimator and blocked estimator. 

To assess the performance of the considered methods, we {use} F-norm to measure the performance of each method.


```{r echo=FALSE, eval = T}

{
  ms = 1:10
  ps = 15 
  es = c(0.02, 0.05, 0.1)
  gammas = seq(0, 10, 2)
}


{
  m = 1
  p = 15
  e = 0.1
  gamma = 0
}


{
  cwf = function(x){covCW(x)$covmatrix}
  mcdf = function(x){robustbase::covMcd(x)$cov}
  ogkf = function(x){robustbase::covOGK(x, sigmamu = robustbase::s_Qn)$cov}
  pairf = function(x){robcovsel::covf(x, cor.method = "pair", 
                                      scale.method = "qn", 
                                      pda.method = F)$covmatrix}
  snipEMf = function(x){
    n = dim(x)[1]
    p = dim(x)[2]
    V = matrix(1, nrow = n, ncol = p)
    V[!is.na(match(as.vector(x),boxplot(as.vector(x),plot=FALSE)$out))] = 0
    snipEM::snipEM(X = x, V = V)$S}
  
  Blockf = function(x){
    covBC(x, group = c(rep(1,5), rep(2,5), rep(3,5)))$covmatrix

  }
  
  mtds = list(Cellweight = cwf, 
              MCD = mcdf,
              OGK = ogkf,
              Pair = pairf,
              snipEM = snipEMf,
              Block = Blockf
  )
}

```

```{r}
library(inline)
openblas.set.num.threads <- cfunction( signature(ipt="integer"),
                                       body = 'openblas_set_num_threads(*ipt);',
                                       otherdefs = c ('extern void openblas_set_num_threads(int);'),
                                       libargs = c ('/usr/lib/x86_64-linux-gnu/openblas-pthread/libblas.so'),
                                       language = "C",
                                       convention = ".C"
)
openblas.set.num.threads(1)
```





```{r echo=FALSE, eval = T}

library(doParallel)
registerDoParallel(cores=8)
#getDoParWorkers()

# parallel computation loops
version = 0
{
  systemtime = system.time({
    result <- foreach(m = ms, 
                      .packages = c("robustbase","robcovsel", "robcovcell"))%:%
      foreach(p = ps)%:%
      foreach(e = es)%:%
      foreach(gamma = gammas)%dopar% {
        
        seed = which(gammas==gamma) + 10*which(es==e) +
          100*which(ps==p) + 1000*which(ms==m) + version 
        set.seed(seed = seed)
    
        dataset = robcovcell::genevar(p = p, e = e, gamma = gamma)
        x = dataset$x
 
        rst = list(cw = NULL, 
                   mcd = NULL, 
                   ogk = NULL, 
                   pair = NULL,
                   snipem = NULL,
                   block = NULL)
        for (mtd in 1:length(mtds)) {
          covmatrix = mtds[[mtd]](x)
          Fnorm = matrixcalc::frobenius.norm(covmatrix-sigma)
          rst[[mtd]] = c(m = m, n = 100, p = p, e = e, gamma = gamma, 
                           mtd = names(mtds)[mtd], seed =seed,  
                           Fnorm = Fnorm)
        }
        rst
      }
  })
  
  
  
  #save(result, file = "result_simu_p20.RData")
}
```

```{r echo=FALSE, fig.width=10, eval=T} 
library(tidyverse)
result1 = as.data.frame(t(as.data.frame(result)))
#glimpse(result1)
result1$Fnorm = as.numeric(result1$Fnorm)
result1$gamma = as.numeric(result1$gamma)

summary = result1 %>% group_by(n,p,e,gamma,mtd) %>% summarise(
  F_norm = mean(Fnorm)) %>% ungroup()
summary = summary %>% mutate(Method = factor(mtd, levels = c("MCD",  "OGK", "Pair","snipEM", "Cellweight","Block")))
summary$e = factor(summary$e, labels = c("e = 2%","e = 5%","e = 10%"))
#glimpse(summary)


ggplot(data = summary, aes(x = gamma,y = F_norm, color = Method ))+geom_line()+geom_point()+facet_wrap(~e) +
  scale_x_continuous(breaks=seq(0,12,2)) +
  ylim(0,20) +
  labs(fill = "", x = expression(gamma*": magnitude of outlyingness"), y = expression(Fnorm))+ 
    theme(text= element_text(size=10))


```


