---
title: "Simulation with structured outlying cells"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, 
  message = F, 
  warning = F , 
  comment = "#>"
)
```

```{r echo=FALSE}
library(cellWise)
library(snipEM)
library(robcovcell)
#remotes::install_github('cran/snipEM')
```

In our simulations, we set $n = 100$, $p = 10$, $\mathbf{x}_i$ is sampled from $N(\pmb 0,\pmb\Sigma_{xx})$. We consider two kinds of covariance matrices $\pmb\Sigma_{xx}$. Type ALYZ are the randomly generated covariance matrices of Agostinelli et al. (2015)
8which typically have relatively small correlations. Type A09 is given by $\Sigma _{ij} = \rho^{|i-j|}$ and we set $\rho = -0.9$. 
Contamination proportions are set as $0\%$, $10\%$ and $20\%$ for all variables separately. Outlying cells of all variables are randomly generated as introduced in Raymaekers and Rousseeuw (2019).

We compared 5 different methods: Gaussrank estimator, GK pairwise estimator (Gnanadesikan and Kettenring, 1972), cellwise weighted estimator, snipEM estimator (Farcomeni 2014) and DDC estimator (Raymaekers and Rousseeuw 2019). 

To assess the performance of the considered methods, we {use} F-norm and KL-divergence to measure the performance of each method.

$$
KL = \frac { 1 } { 2 } ( \operatorname { tr } ( \hat\Sigma  ^ { - 1 } \Sigma _ { 0 } ) - p + \operatorname { ln } ( \frac { \operatorname { det } \hat\Sigma } { \operatorname { det } \Sigma _ { 0 } } ) )
$$


```{r echo=FALSE, eval = T}

{
  m = 1
  p = 10
  e = 0.1
  gamma = 10
  sigmatype = "A09"
}

snipEMf = function(x){
  n = dim(x)[1]
  p = dim(x)[2]
  V = apply(x,2, function(xvec){as.numeric(is.na(match(xvec,boxplot(xvec,plot=FALSE)$out)))} )
}

Grankf = function(x){
  robcovsel::covf(x, cor.method = "gaussrank", scale.method = "qn", pda.method = F)$cormatrix
}

pairf = function(x){robcovsel::covf(x, cor.method = "pair", 
                                    scale.method = "qn", 
                                    pda.method = "nearpd")$cormatrix}


mtds = list(Grank = Grankf,
            Pair = pairf,
            snipEM = snipEMf
)

{
  Grankf = function(x){
    robcovsel::covf(x, cor.method = "gaussrank", scale.method = "qn", pda.method = F)$covmatrix
  }
  pairf = function(x){robcovsel::covf(x, cor.method = "pair", 
                                      scale.method = "qn", 
                                      pda.method = "nearpd")$covmatrix}
  snipEMf = function(x){
    n = dim(x)[1]
    p = dim(x)[2]
    V = apply(x,2, function(xvec){as.numeric( robustbase::covMcd(xvec)$mcd.wt )} )
    
    snipEM::snipEM(X = x, V = V)$S}
  
  cwf = function(x,bi){robcovcell::covCW(x)$covmatrix}
  
  DDCf = function(x, bi){
    fit = cellWise::DDC(x)
    cov(fit$Ximp) 
    }
  

  
  mtds = list(Grank = Grankf,
              Pair = pairf,
              snipEM = snipEMf,
              Cellweight = cwf,
              DDC = DDCf
              
  )
}

```




```{r echo = F,  eval=T}
library(inline)
openblas.set.num.threads <- cfunction( signature(ipt="integer"),
                                       body = 'openblas_set_num_threads(*ipt);',
                                       otherdefs = c ('extern void openblas_set_num_threads(int);'),
                                       libargs = c ('/usr/lib/x86_64-linux-gnu/openblas-pthread/libblas.so'),
                                       language = "C",
                                       convention = ".C"
)
openblas.set.num.threads(1)
```





```{r echo=FALSE, message = F, eval = T}

{
  ms = 1:100
  ps = 10
  es = c(0, 0.1, 0.2)
  gammas = seq(0, 10, 1)
  sigmatypes = c("ALYZ", "A09")
}

library(doParallel)
registerDoParallel(cores=8)
#getDoParWorkers()


# parallel computation loops
{
  systemtime = system.time({
    result <- foreach(m = ms, 
                      .packages = c("robustbase","robcovsel", "robcovcell", "cellWise", "snipEM"))%:%
      foreach(p = ps)%:%
      foreach(e = es)%:%
      foreach(sigmatype = sigmatypes)%:%
      foreach(mtd = 1:length(mtds))%:%
      foreach(gamma = gammas)%dopar% {
        
        set.seed(seed = m)
        n=100
        mu = rep(0,p)
        Sigma = cellWise::generateCorMat(d = p, corrType = sigmatype)
        dataset = cellWise::generateData(n = n, d = p, mu = mu, Sigma = Sigma, perout = e,
                      gamma = gamma, outlierType = "cellwiseStructured")
        x = dataset$X
        indcells = dataset$indcells
        rst = list(grank = NULL,
                   pair = NULL,
                   snipem = NULL,
                   DDC = NULL
                   )
        for (mtd in 1:length(mtds)) {
          covmatrix = mtds[[mtd]](x)
          Fnorm = matrixcalc::frobenius.norm(covmatrix-Sigma)
          KLd = robcovcell::KLdiv(xhat = covmatrix, x0 = Sigma)
          rst[[mtd]] = c(m = m, n = 100, p = p, e = e, gamma = gamma, sigmatype = sigmatype,
                           mtd = names(mtds)[mtd], Fnorm = Fnorm, KLdiv = KLd)
        }
        rst
      }
  })
  
  
  #systemtime
  #save(result, file = "result_simu_p20.RData")
}
```

```{r echo=FALSE, eval=T} 
library(tidyverse)
result1 = as.data.frame(t(as.data.frame(result)))
#glimpse(result1)
result1$Fnorm = as.numeric(result1$Fnorm)
result1$gamma = as.numeric(result1$gamma)
result1$KLdiv = as.numeric(result1$KLdiv)

summary = result1 %>% group_by(n,p,e,gamma,sigmatype, mtd) %>% summarise(
  F_norm = mean(Fnorm, na.rm = T), KL_div = mean(KLdiv, na.rm = T)) %>% ungroup()
summary = summary %>% mutate(Method = factor(mtd, levels = c("Grank","Pearson","MCD",  "OGK", "Pair","snipEM", "snipEM2", "Cellweight", "Cellweight2", "DI", "DDC")))
summary$e = factor(summary$e, labels = c("e = 0%",#"e = 5%",
                                         "e = 10%","e = 20%"))
summary$sigmatype = factor(summary$sigmatype, levels = c("ALYZ", "A09"))
#glimpse(summary)

```



```{r echo=FALSE, fig.width=6, fig.height=6, eval=T}
hist(result1$KLdiv)
```


```{r echo=FALSE, fig.width=12, fig.height=8, eval=T}

p1 = ggplot(data = summary, aes(x = gamma,y = F_norm, color = Method ))+geom_line()+geom_point()+
  facet_grid(sigmatype~e ) +
  scale_x_continuous(breaks=seq(0,12,2)) +
  #ylim(0,5) +
  labs(fill = "", x = expression(gamma*": magnitude of outlyingness"), y = expression(Fnorm))+ 
    theme(text= element_text(size=10))

p1


```

```{r echo=FALSE, fig.width=12, fig.height=8, eval=T} 
p2 = ggplot(data = summary, aes(x = gamma,y = KL_div, color = Method ))+geom_line()+geom_point()+
  facet_grid(sigmatype~e)+
  scale_x_continuous(breaks=seq(0,12,2)) +
  #ylim(0,150) +
  labs(fill = "", x = expression(gamma*": magnitude of outlyingness"), y = expression(KLdiv))+ 
    theme(text= element_text(size=10))

p2


```


